An Efficient Feedback Calibration Algorithm for Direct Imaging Radio Telescopes

### 2016-05-03 cal_demo/ms.tex
Starting to add text to explain FX and clarify noise assumptions

The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
components of the electric field we already discussed. The sampling rate of the digitization 
determines the bandwidth of the observation, and the sampling period determines the minimum 
frequency channel resolution. The frequency Fourier transform operation is often referred to as 
the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross-
correlator which performs the frequency transform (F-engine) before cross multiplying (X-
engine).

@@

Thus far we have assumed the antennas deliver the true electric field signals. Here we 
introduce a multiplicative complex gain as well as an additive noise term which corrupt the 
measured signals.
\begin{equation}\label{eq:apply_gain}
\Er{a}(f,t) = g^T_a(f,t) \Er{a}^T(f,t) + \widetilde{n}_a(f,t)
\end{equation}
Note that this quantity is neither a true or estimated value, but rather the raw electric field signal 
corrupted by the complex gain before we attempt to estimate the true value. The noise term is 
strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
in the time dependence of the sky electric field. We will neglect $\widetilde{n}_a$ for now, but 
will inspect its effects at the end of this section. Each subsequent step treats each frequency 
channel independently, so we will drop the $f$ to simplify notation.

The MOFF algorithm next calls for a calibration. The goal of our new calibration method will be 
to form an estimate of the gains, $g'_a$. For now we will assume we have formed an estimate 
to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
current estimate of the gains.
\begin{equation}
\Erest{a}(t) = \Er{a}(t)/ g'_a
\end{equation}

The next step is to grid the antenna signals. MOFF uses the antenna patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. This 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
operation must be performed for every time step, before any averaging of the data.

>>>>

The first step of the MOFF algorithm is to digitize the antenna signals and perform an FFT to 
channelize the time-ordered data. This is effectively estimating the quasi-monochromatic 
components of the electric field we already discussed. The sampling rate of the digitization 
determines the bandwidth of the observation, and the sampling period determines the minimum 
frequency channel resolution. The frequency Fourier transform operation is often referred to as 
the F-engine of a correlator. The MOFF shares the same F-engine design as an FX cross
correlator which performs the frequency transform (F-engine) before cross multiplying 
(X-engine).

@@

Thus far we have assumed the antennas perfectly measure the sky electric field distribution according to their far-field radiation patterns. Here we introduce a multiplicative complex gain as well as an additive noise term which corrupt the measured signals.
\begin{equation}\label{eq:apply_gain}
E_a(f,t) = g_a(f,t) E_a^I(f,t) + n_a(f,t)
\end{equation}
The noise term is 
strictly receiver noise -- noise introduced by the instrument. Any sky noise is implicitly included 
in the time dependence of the sky electric field. 
We have carried the frequency dependence to this point to make clear the complex gains are
frequency dependent. However, each subsequent step treats each frequency 
channel independently, so we will drop the $f$ to simplify notation. 
Furthermore, while antenna gains may vary with time, we will assume they are constant over
the period of our calibration, and thus drop their time dependence from the notation.

The MOFF algorithm next calls for a calibration of the electric field measurements. The goal of our new calibration method will be 
to form an estimate of the gains, $g'_a$, where the tick represents the estimate. For now we will assume we have formed an estimate 
to proceed with the MOFF pipeline. We correct the incoming electric field data stream using our 
current estimate of the gains.
\begin{equation}
E'_a(t) = E_a(t)/ g'_a
\end{equation}

The next step is to grid the calibrated electric field measurements. MOFF uses the antenna patterns as the gridding 
kernel according to the software holography/A-transpose technique \citep{mor09,bha08}. 
Mathematically we can express the gridding operation as
\begin{equation}
E'_j=\sum_a W_a(\mathbf{r}_i-\mathbf{r}_a) E'_a(t)
\end{equation}
This 
achieves an optimal map \citep{teg97b} and places the data on a regular grid. Unlike 
traditional correlators which grid visibilities, the MOFF grids the electric fields directly. This 
operation must be performed for every time step, before any averaging of the data.
